# Introduction {#intro}

The machine learning algorithms were trained to detect breast
cancer using the Wisconsin Diagnostic Breast Cancer (WDBC)
dataset[20]. The dataset consists of features which
were computed from a digitized image of a fine needle aspirate
(FNA) of a breast mass. The said features describe the characteristics
of the cell nuclei found in the image.


## Description of the data


```{r}
dat <- read.csv("data/WBCD_data.csv")
knitr::kable(head(dat))
```



## Check how may variables are there. 

```{r}
dim(dat)
```


As you can see, there are `r nrow(dat)` rows and `r ncol(dat)` columns. 

## Check the colum names

```{r}
colnames(dat)
```

The name of last column is `X` is because the name is not provided in the data. The first column is the `ID` of the individuals, the second column `diagnosis` is a categorical variable. 


```{sh}
head data/WBCD_data.csv
```


## 

```{r}
str(dat)
```

## Remove last column 

```{r}
dim(dat)
dat$X=NULL 
dim(dat)
```

```{r}
dat_predictor = dat[, 3:ncol(dat)]
```

Now let's only extract the predictor variables. 



## Check the correlation among the predictor variables. 


```{r}
cov_matrix = cor(dat_predictor)
```

```{r}
knitr::kable(cov_matrix)
```


```{r}
library(corrplot)
corrplot(cov_matrix, type="upper", order="hclust", tl.col="black", tl.srt=45, tl.cex = 0.6)
```


## Identifying Correlated Predictors

```{r}
summary(cov_matrix[upper.tri(cov_matrix)])
```

```{r}
highCorr <- sum(abs(cov_matrix[upper.tri(cov_matrix)]) > .9)
highCorr
```


```{r}
highlyCorDescr <- findCorrelation(cov_matrix, cutoff = .75)
filteredDescr <- dat_predictor[,-highlyCorDescr]
descrCor2 <- cor(filteredDescr)
summary(descrCor2[upper.tri(descrCor2)])
```


```{r}
filteredDescr <- filteredDescr[,-highlyCorDescr]
```





